Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-09-20 documents/mixture_models.tex
objective function and optimization

...If we wanted to approximate these models, what objective function
would we use?  What tolerances are okay?...Define a sensible scalar
objective...Give fitting procedure...give actual values for the
parameters...

@@

@@

Applying this PSF to an exponential or de~Vaucouleurs galaxy is
slightly more complicated, because the galaxy has not just a flux
$S_g$ and a central position $\pos_g$; it also has a shape.  Because
we are only considering these simple galaxies, we are only permitting
ellipsoidal shapes, which can be represented by a semi-major axis $a$,
a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or by a symmetric
variance tensor $\var_g$ (which is a general representation of an
ellipse).  The galaxy is distorted by this variance tensor
\emph{prior} to PSF convolution, so the focal-plane image
(PSF-convolved intensity field) for a general exponential galaxy is
given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_{\exp}} S_g\,a_{\exp,m}\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var_{\exp,m}\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors on the sky pointing in the major-axis and minor-axis
directions respectively.  Implicitly all vectors are two-dimensional
column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
matrix that contains the ``shape'' (position angle, major-axis, and
ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
$\lux$ cases are all similar.  Note the important and key result of
this \documentname, to wit, that a mixture-of-Gaussian galaxy model
(with $M$ components) convolved with a mixture-of-Gaussian PSF model
(with $K$ components) yields a mixture-of-Gaussian image (with
$[M\,K]$ components).

@@

...what would you want to do to generalize to S\'ersic profiles?...

@@

\acknowledgements
It is a pleasure to thank...

>>>>

We seek the best mixture-of-Gaussian approximations.  This
necessitates definition of the word ``best''.  If we think of the
profiles as being two-diensional probability distribution functions
(for, say, the arrivals of photons), then one natural choice is the
K-L divergence or similar cross-entropy or information-theoretic
measure.  However, in typical astronomical imaging, the galaxy is
superimposed on a substantial, flat sky level, and the noise in the
data is close to Gaussian.  This suggests more chi-squared-like
objectives.  We adopt the latter, in part because it is most
appropriate for our specific proposed application (modeling \sdss-like
astronomical imaging), but experiments we have performed suggest that
information-theoretic objectives also lead to good results.

In detail, the chi-squared objective we minimize---the
\emph{badness}---is a squared residual between function and
approximation.  It is designed to be equivalent to a chi-squared
statistic in a homoskedastic two-dimensional image of the profile
taken with extremely high angular resolution (pixels of size 0.001 the
half-light radius) and vanishing point-spread function.  The badness
is computed in one dimension but weighted in radius (weight increasing
linearly with radius) to make it equivalent to the two-dimensional
chi-squared.  Optimization is performed by the \project{scipy}
implementation of the BFGS algorithm, with many initializations to
explore multiple local minima.  Further details are available in the
code, which is publicly
available.\footnote{\url{https://github.com/davidwhogg/TheTractor/}}

@@

...give tables of results, including Bundy versions and recommended versions...

...show dependence of amplitudes and variances on sersic n for some M...

@@

Applying this PSF to an exponential or de~Vaucouleurs galaxy is
slightly more complicated, because the galaxy has not just a flux
$S_g$ and a central position $\pos_g$; it also has a shape.  Because
we are only considering these simple galaxies, we are only permitting
ellipsoidal shapes, which can be represented by a semi-major axis $a$,
a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
an affine transformation $\affine_g$ that takes a circle to the
relevant ellipse (and is a therefore a general representation of an
ellipse; it is also the matrix square root of the symmetric variance
tensor describing the ellipse).  The galaxy is distorted by this
affine transformation \emph{prior} to PSF convolution, so the
focal-plane image (PSF-convolved intensity field) for a general (say)
$\exp$ galaxy is given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_{\exp}} S_g\,a_{\exp,m}\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var_{\exp,m}\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors on the sky pointing in the major-axis and minor-axis
directions respectively.  Implicitly all vectors are two-dimensional
column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
matrix that contains the ``shape'' (position angle, major-axis, and
ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
$\lux$ cases are all similar.  Note the important and key result of
this \documentname, to wit, that a mixture-of-Gaussian galaxy model
(with $M$ components) convolved with a mixture-of-Gaussian PSF model
(with $K$ components) yields a mixture-of-Gaussian image (with
$[M\,K]$ components).

@@

@@

\acknowledgements It is a pleasure to thank Brendon Brewer (Auckland)
and Kevin Bundy (IPMU) for valuable comments.  This work was supported
in part by NASA (grant NNX12AI50G) and the NSF (grant IIS-1124794).
